Quantum H theorem, variational entropy, and relaxation processes

### 2021-01-22 tex/disscussion.tex
se agregaron los ultimos cambios, se unió la sección de discusión con conclusiones. La sección de relajación no se ha modificado hasta haber discutido lo que propone el Dr. Carrillo.

\section{Discussions and remarks}\label{sec:disscussion}

In the literature, the Boltzmann's $H$-theorem is proved using a spatially homogeneous distribution function. In this work, we used the local equilibrium hypothesis and as a consequence of this, it was necessary to include the homogeneous distribution function hypothesis. However, we included inhomogeneities in the system through the division of the volume $V$ in cells and the non-homogeneous distribution function assumption in the total volume. Moreover, the Boltzmann's $H$-theorem still holds when the system includes inhomogeneities, only if the local equilibrium hypothesis holds. We can include the fact that we proved the Boltzmann's $H$-theorem without using the Boltzmann's transport equation. It was only necessary to specify the behavior of the deviation $g$ and its derivative in time. Also, we obtained the Boltzmann's transport equation in equilibrium from the Boltzmann's $H$-functional through the variational method. Nevertheless, in order to obtain the complete Boltzmann's transport equation, it was necessary to include the collision term because of the variation of the $H$-functional is not zero when the system is out of equilibrium. This is a consequence of the $H$-theorem.\\
\\
In the quantum version, we used the same assumptions to prove the $H$-theorem, those were the behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the non-homogeneous distribution function assumption. With the proof of the quantum version of the $H$-theorem, we can say that a quantum gas inside a volume $V$ that is out of equilibrium with inhomogeneities, in the first-order approximation the system evolves to the equilibrium state as long as the local equilibrium hypothesis holds. In addition to this, the system is in the equilibrium state when the derivative of the variational entropy with respect to time is equal to zero, and as a consequence of this, the distribution function corresponds to the Bose-Einstein or Fermi-Dirac distribution.\\
On the other hand, the equation of motion obtained from the variational entropy could be seen as a quantum transport equation. However, in order to develop the transport phenomena (in analogy to theoretical work developed by Boltzmann) is necessary to find an expression for the collision term. This problem may be future work.\\
\\
We also find that variational entropy is proportional to the Boltzmann's $H$-functional in the limit of high energy levels. This conclusion is trivial if we compare (\ref{h-quantic4}) with (\ref{reduce-h}). Moreover, if we see the expression for the Boltzmann's $H$-functional in equilibrium
\begin{equation}
    H_{B}=-\frac{S}{Vk},
\end{equation}
where $S$ is the entropy of the system, and $k$ is the Boltzmann's constant, we can identify that variational entropy is proportional to the entropy of the system.\\
\begin{equation}
    \Ss\propto S.
\end{equation}
Besides, we find an expression for the entropy density (\ref{H-entropy}). This expression will be useful to describe quantum gases in terms of entropy.\\
It is important to remark that we find that $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$ setting the expecting behavior of the deviation $g$. This correct behavior involves the molecular chaos hypothesis in the classical case and the random a priori phases hypothesis in the quantum case. Specifically those hypothesis are included in $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$. Those time derivatives correspond to the expected behavior in a meantime, that is, $\dot{g}_{nJ}^{+}$ always decreases and $\dot{g}_{nJ}^{-}$ always increases.\\ 
According to the classical analysis of the $H$-theorem, the expected behavior of the distribution function is presented in the system if the molecular chaos hypothesis holds, but in a meantime, called \textit{relaxation time}, the distribution function behaves as we expect, such as the distribution function evolves to the equilibrium state keeping that $\frac{dH}{dt}\leq 0$.\\
In the same way as the previous statement, we proposed that $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ have the expected behavior in a relaxation time. Also molecular chaos hypothesis and a random a priori phases hypothesis are implicit in the expected behavior of the deviation such as $\frac{dH'}{dt}\leq 0$ and $\frac{d\Ss}{dt}\leq 0$.

>>>>

\section{Remarks and conclusions}\label{sec:disscussion}

In the literature, the Boltzmann's $H$-theorem is proved using a spatially homogeneous distribution function. 
Here, we considered inhomogeneities in the system through the division of the volume $V$ in cells 
and the non-homogeneous distribution function assumption in the total volume. 
We have shown that the Boltzmann's $H$-theorem still holds when the system includes 
inhomogeneities, of course assuming that the local equilibrium hypothesis holds. 
We demostrated the classical Boltzmann's $H$-theorem without using explicity the Boltzmann's transport equation. 
Instead of that, it needed to specify the behavior of the deviation $g$ and its time derivative. 
\\
In the quantum version, we used the same assumptions to prove the $H$-theorem, those were the 
behavior of the deviation $g_{nM}$ and its derivative, the local equilibrium hypothesis, and the 
non-homogeneous distribution function assumption. 
We have corroborated the correctness of the $f$ expression in equilibrium recovering 
the MB, FD and BE distributions and in recovering the expression of the entropy of an ideal gas. 
We demostrated as well that the functional $H$ leads to the proper fulfilling of the quantum classical correspondence.
\\
\\
We also have shown that variational entropy is proportional to the Boltzmann's $H$-functional in the non-degenerated limit. 
This is a straightfowars conclusion. One may infer from the expression for the Boltzmann's $H$-functional in equilibrium that
\begin{equation}
    H_{B}=-\frac{S}{Vk},
\end{equation}
where $S$ is the entropy of the system, and $k$ is the Boltzmann's constant, we can identify that variational entropy is proportional to the entropy of the system.\\
\begin{equation}
    \Ss\propto S.
\end{equation}
It is important to remark that we find that $\frac{d\mathcal{H}'}{dt}\leq 0$ and $\frac{d\mathcal{H}}{dt}\leq 0$ 
setting the expecting behavior of the deviation $g$. 
This correct behavior results from the molecular chaos hypothesis in the classical case and 
the random a priori phases hypothesis in the quantum case. 
Specifically those hypothesis are included in the assumed behavior of $\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$. 
Those time derivatives correspond to the expected behavior in a meantime, that 
is, $\dot{g}_{nJ}^{+}$ always decreases and $\dot{g}_{nJ}^{-}$ always increases.\\ 
According to the classical $H$-theorem, if the molecular chaos hypothesis holds, 
although within small times of the order of the \textit{relaxation time}, as a consequence of the 
stochastic nature of the collisions, $H$ function evolves through a sequential series of 
molecular chaos and non-molecular chaos, but a larger times $\frac{dH}{dt}\leq 0$.\\
In analogy with this, we see that in our procedure this behavior is introduced by 
$\dot{g}_{nJ}^{+}$ and $\dot{g}_{nJ}^{-}$ to obtain the expected behavior in a relaxation time. 
So, molecular chaos hypothesis and a random a priori phases hypothesis are implicit
 in the expected behavior of the deviation from the equilibrium of the $H$ function such that 
$\frac{d\mathcal{H}'}{dt}\leq 0$ and $\frac{d\mathcal{H}}{dt}\leq 0$.
\\
\\
In summaty, we propose a variational procedure to demostrate the classical and quantum $H$ theorem, that
make possible to describe at a mesoscopic local view (cell-scale), the time evolution 
of an out-of-equilibrium, spatially non-homogeneous, system to the equilibrium condition.
In principle, this approach would make possible investigate the transport phenomena inherent to the
equilibrium process, starting from an arbitrarily far-from-equilibrium initial condition.

### 2021-01-22 tex/qccorrespondence.tex
se agregaron los ultimos cambios, se unió la sección de discusión con conclusiones. La sección de relajación no se ha modificado hasta haber discutido lo que propone el Dr. Carrillo.

We start by arguing that, in the classic limit, $f_{nM}\approx0$ for the following reasons.
Systems at very low temperatures, where the quantum effects cannot be ignored, are obviously excluded
from the current analysis. For systems at sufficently large temperature, the particles
occupy almost exclusively high-energy levels. 
Furthermore, the energy spectrum approaches to be a continuum, as is expected by taking the limit
$\hbar\to0$, and the number of particles per level are very close to zero.

@@

Because we have shown that both classical and quantum $H$-functionals, Eqs.~(\ref{eq:cHdef}) and
(\ref{eq:qHdef}) recover the Maxwell-Boltzmann, Bose-Einstein, or Fermi-Dirac distribution
functions in equilibrium, we will omit the quantum classical correspondence of the distribution functions,
as it has been thoroughly discussed in textbooks.

>>>>

We start by arguing that, in equilibrium, it is a trivial task to proof that expression (\ref{eq:qfneqtwo}) collapses
into expression (\ref{eq:fbardef}) by simply letting that the 
degeneration parameter $\xi=e^{-\frac{1}{K_B T}(\epsilon-\mu)} \gg 1$.
A more general approach to see the quantum classical correspondence would be if the quantum functional 
$H$ collapsed  into the classical one in the appropiated limit.
This limit in our case will be $f_{nM} \approx 0$ because the following reasons.
Systems at very low temperatures, where the quantum effects cannot be ignored, are obviously excluded
from the current analysis. For systems at sufficently large temperature, the particles
occupy almost exclusively high-energy levels. 
Furthermore, the energy spectrum approaches to be a continuum, as is expected by taking the limit
$\hbar\to0$, and the number of particles per level are very close to zero.

@@
